---
title: "Bank Product Recommendations - Complete Model without Evaluation"
author: Grewal, Dinesh (imdineshgrewal@gmail.com)
output: html_document
---

###Introduction
An algorithm which helps bank to predict which products their existing customers will use in the next month based on their past behavior.

####Data Explore

#####Libraries

```{r results='hide'}
#install.packages("rmarkdown")
#install.packages("knitr")
#install.packages("shiny")
#install.packages("readr")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("tidyr")
library(knitr)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(data.table)
```


#####Load Data from CSV files :
```{r results='hide'}
data_raw <- read_csv("~/data.csv")
```


#####Data Dimension :
```{r}
dim(data_raw)
```


#####Data summary :
```{r}
summary(data_raw)
```


#####variables(column) names :
```{r}
names(data_raw)
```


#####Column Rename :
```{r results='hide'}
setnames(data_raw, old=c("fecha_dato", 
                         "ncodpers", 
                         "ind_empleado", 
                         "pais_residencia",
                         "sexo",
                         "fecha_alta",
                         "ind_nuevo",
                         "antiguedad",
                         "indrel",                
                         "ult_fec_cli_1t",
                         "indrel_1mes",
                         "tiprel_1mes",           
                         "indresi",               
                         "indext",                
                         "conyuemp",    
                         "canal_entrada",         
                         "indfall",
                         "tipodom",
                         "cod_prov", 
                         "nomprov",
                         "ind_actividad_cliente", 
                         "renta",
                         "ind_ahor_fin_ult1",     
                         "ind_aval_fin_ult1",    
                         "ind_cco_fin_ult1",      
                         "ind_cder_fin_ult1",     
                         "ind_cno_fin_ult1",      
                         "ind_ctju_fin_ult1",    
                         "ind_ctma_fin_ult1",     
                         "ind_ctop_fin_ult1",     
                         "ind_ctpp_fin_ult1",     
                         "ind_deco_fin_ult1",    
                         "ind_deme_fin_ult1",     
                         "ind_dela_fin_ult1",     
                         "ind_ecue_fin_ult1",     
                         "ind_fond_fin_ult1",    
                         "ind_hip_fin_ult1",     
                         "ind_plan_fin_ult1",     
                         "ind_pres_fin_ult1",     
                         "ind_reca_fin_ult1",    
                         "ind_tjcr_fin_ult1",     
                         "ind_valo_fin_ult1",     
                         "ind_viv_fin_ult1",      
                         "ind_nomina_ult1",      
                         "ind_nom_pens_ult1",     
                         "ind_recibo_ult1"
), 
new=c("date", 
      "custcode", 
      "emp_index", 
      "country",
      "gender",
      "first_date",
      "new_customer",
      "senior",            
      "primary",            
      "last_date",
      "customer_typ",
      "customer_rel_typ",
      "res_indx",
      "for_indx",
      "spouse_indx",
      "channel",
      "dec_indx",
      "add_type",
      "prov_code",
      "prov_name",
      "cust_active",
      "income",
      "saving_Acc",
      "guarantees",
      "curr_Acc",
      "deri_Acc",
      "payroll_acc",
      "junior_acc",
      "Más_part_Acc",
      "part_Acc",
      "part_Plus_Acc",
      "Shrt_term_dep",
      "Mid_term_dep",
      "Long_term_dep",
      "e_acc",
      "funds",
      "mortgage",
      "pensions",
      "loans",
      "taxes",
      "credit_card",
      "securities",
      "home_acc",
      "payroll",
      "nom_pensions",
      "direct_debit"
))
```


#####New Column Names :
```{r}
names(data_raw)
```


#####Glimpse of Data through tidyr package :
```{r}
glimpse(data_raw)
```


#####Check if there any NA's present in the data :
```{r}
summary(factor(sapply(data_raw, function(x)any(is.na(x)))))
sapply(data_raw, function(x)any(is.na(x)))
```


####Data Cleaning
#####Employee Index
*emp_index
-Employee index: A active, B ex employed, F filial, N not employee, P pasive.
-Remove record that having missing emp_index.
```{r}
data_raw <- subset(data_raw, data_raw$emp_index!="")
```


#####Date format correction :
```{r}
data_raw$date <- as.POSIXct(strptime(data_raw$date,format="%Y-%m-%d"))
data_raw$first_date <- as.POSIXct(strptime(data_raw$first_date,format="%Y-%m-%d"))
```


#####Numeric data type conversion :
```{r}
data_raw$age <- as.numeric(data_raw$age)
data_raw$income <- as.numeric(data_raw$income)
```


#####Age
* Data Filtering with age
- filter data with less then 100 year people with graph.
```{r}
data_raw <- subset(data_raw, data_raw$age>=18 & data_raw$age<=100)
hist(data_raw$age)
```

People with age 20-25 having higest frequeny and people from 25-20 having average frequency
60-100 decresing low.


#####Gender
*Data Filtering with Gender
-Removing NA from Gender 
```{r}
summary(factor(data_raw$gender))
data_raw  <- subset(data_raw, !is.na(gender))
summary(factor(data_raw$gender))
```


#####segmentation
* segmento
- 01 - VIP, 02 - Individuals 03 - college graduated
- Removing NA's (Subset)
```{r}
summary(factor(data_raw$segmento))
#data_raw  <- subset(data_raw, !is.na(segmento))
```


#####channel used by the customer to join
* canal_entrada       
- Channel used by the customer to join
- Removing NA's (Subset)
```{r}
summary(is.na(data_raw$channel))
#Mode    FALSE       TRUE 
#logical 13326623    847

data_raw  <- subset(data_raw, !is.na(channel))
```


#####Salary
* Salary Distibution
- Income NA values -> Median.
```{r}
summary(data_raw$income)
summary(factor(is.na(data_raw$income)))
data_raw$income[is.na(data_raw$income)] <- median(data_raw$income, na.rm=TRUE)
####
#count(filter(data_raw, income >= 4000000)) #upper limit
#count(filter(data_raw, income <= 0)) #low limit

#df <- filter(data_raw, income >= 28894390)
#df$nomprov

#ggplot(data_raw, aes(x=nomprov, y=income)) + geom_point()
```

MADRID prov_name is having the highest salary 8894396


#####Province
* Province name
- Data is important, even for those who's place is unknown.
- Random string and value 0 in the _code.
```{r}
summary(factor(data_raw$prov_name))
data_raw$prov_name[is.na(data_raw$prov_name)==TRUE] <- 'unknown'
data_raw$prov_code[is.na(data_raw$prov_code)==TRUE] <- 0
summary(factor(is.na(data_raw$prov_name)))
```


#####Customer Type
* customer_typ
- Customer type at the beginning of the month.
- 1 (First/Primary customer), 2 (co-owner ),P (Potential),3 (former primary), 4(former co-owner).
```{r}
summary(factor(data_raw$customer_typ))
#1            2        3         4     NA's 
#13359714     1317     4338      305   122183 
data_raw$customer_typ[is.na(data_raw$customer_typ)] <- 1
```


#####Payroll
* ind_nomina_ult1 
- Highest variable values to NA's
```{r}
summary(factor(data_raw$payroll))
#       0        1     NA's 
#12581324   745165      134 
data_raw$payroll[is.na(data_raw$payroll)] <- 0
```


#####Pensions
* ind_plan_fin_ult1 
- Highest variable values to NA's
```{r}
summary(factor(data_raw$pensions))
#       0        1     NA's 
#12517530   808959      134 
data_raw$pensions[is.na(data_raw$pensions)] <- 0
```


#####Nom Pensions
* ind_nom_pens_ult1
- Highest variable values to NA's.
```{r}
summary(factor(data_raw$nom_pensions))
#0        1     NA's 
#12521577   808972      149 
data_raw$nom_pensions[is.na(data_raw$nom_pensions)] <- 0
```

#####Customer Relation Type
* customer_rel_typ
- Customer relation type at the beginning of the month, 
- A (active), I (inactive), P (former customer),R (Potential)
- NA to highest value
```{r}
summary(factor(data_raw$customer_rel_typ))
data_raw$customer_rel_typ[is.na(data_raw$customer_rel_typ)] <- 'I'
```


#####spouse_indx & The date in which the customer became as the first holder of a contract in the bank
```{r}
summary(factor(data_raw$spouse_indx))
summary(factor(data_raw$last_date))
```

#####Cleaned Data

####Exploratory Data Analysis
#####Data Visualization

```{r}
#install.packages("lubridate")
library(lubridate)
#install.packages("ggthemes")
library(ggthemes)
library(ggplot2)
#install.packages("corrplot")
library(corrplot)
```


#####Records from cleaned Data Set
```{r}
head(data_raw)
```


#####Clean Data Import
```{r}
clean_data <- data_raw
```


#####Product related data
```{r}
product <- data_raw[,25:48]
```


####Data Exploration
#####1. Exploring(Distribution) Customer Age 
#####Variable converstion form integer to numeric
```{r}
class(clean_data$age)
clean_data$age<-as.numeric(as.character(clean_data$age))
```

#####Visualization
```{r}
ggplot(clean_data, aes(x=age)) +
  stat_count(width=1, position="stack") +
  ggtitle("Age Histogram")
```

#####2. age VS segmento (01 - VIP, 02 - Individuals 03 - college graduated)
#####Segmento into factor
#####clean_data$segmento <- factor(clean_data$segmento, labels = c("VIP", "Individuals","Graduates"))
#####Visualization (age VS segmento)
```{r}
ggplot(clean_data, aes(x=age, fill=factor(segmento))) +
  geom_bar() +
  facet_grid(".~segmento") + 
  scale_fill_discrete(name="Segmantation") +
  theme(legend.position="top") +
  scale_y_continuous("Frequency") 
```



#####3. gender VS segmento
#####Segmento into factor
#####Gender 
```{r}
clean_data$gender <- factor(clean_data$gender, labels = c("Male", 
                                                          "Female"))
```


#####Visualization (gender VS segmento)
```{r}
ggplot(clean_data, aes(x=gender,  fill = factor(gender))) +
  geom_bar() +
  facet_grid(".~segmento") + 
  scale_fill_discrete(name="Gender") +
  theme(legend.position="top") +
  scale_y_continuous("Frequency") 
```


#####4 Changing Age Distribution Over Years(opening accounts)
#####first_date(Year) VS Age
#####COnverstion of the date into years
```{r}
clean_data %>% 
  group_by(year=floor_date(first_date, "year")) %>%
  summarize(age=mean(age)) %>%
  ggplot(aes(x=year, y= age)) + 
  geom_line() +
  xlab("Year") + ylab("Average Age")
```


#####5  Exploring Customer Income
```{r}
ggplot(clean_data, aes(x=segmento ,y=income)) +
  geom_point() +
  # facet_grid(".~segmento") + 
  scale_fill_discrete(name="Segmentation") 
# theme(legend.position="top") 
```

#####6 Product correlation with each other 
```{r}
product_corr <- cor(product)
corrplot(product_corr, method="square")
```

#####Correlation on Graph
```{r}
layout(matrix(1:1, ncol = 3))
corrplot(product_corr, method="number", tl.cex = 1)
```


#####7 Most popular products in customers.
```{r}
product_tbl<-colSums(product)
barplot(product_tbl,
        main = "Most Popular Product",
        col = "darkred",
        las=2,
        cex.axis=0.7,
        cex.names=0.7)
```


#####Sorted all top product
```{r}
all_prod_name <- head(sort(product_tbl,decreasing=TRUE), n = 24)
```

#####Graph
```{r}
barplot(all_prod_name,
        main = "Most Popular Product",
        col = "darkred",
        las=2,
        cex.axis=0.7,
        cex.names=0.7)
```


#####TOP Most product
```{r}
top_prod_name <- head(sort(product_tbl,decreasing=TRUE), n = 5)
top_prod_name
```


#####Data with top 5 products
```{r}
data_top_product<-cbind(clean_data[,1:24], 
                        clean_data[,27], #curr_Acc 
                        clean_data[,29], #payroll_acc
                        clean_data[,32], #part_Acc
                        clean_data[,37], #e_acc
                        clean_data[,48]) #direct_debit
names(data_top_product)
```


####Model-Collaborative Filtering (Item Based)
#####RECOMMENDATION SYSTEM

```{r}
#install.packages("arules")
library(arules)
```

```{r}
dim(data_top_product)
```


#####Select Active customer in the bank
```{r}
data_top_product_active <- subset(data_top_product, data_top_product$cust_active==1)
dim(data_top_product_active)
names(data_top_product_active)
```


#####Data Matrix
```{r}
data.matrix <- data_top_product_active
```


#####Custumer ID to Character
```{r}
data.matrix$custcode <- as.character(data.matrix$custcode)
```


#####Convert age (Continuous Variable) Into A Categorical Variable
```{r}
data.matrix$age <- discretize(data.matrix$age, method = "frequency", 5) 
data.matrix$new_customer <- as.logical(data.matrix$new_customer)
data.matrix$primary <- as.factor(data.matrix$primary)
data.matrix$prov_code<-as.factor(data.matrix$prov_code)
```


#####Convert income (Continuous Variable) Into A Categorical Variable
```{r}
data.matrix$income <- discretize(data.matrix$income, method = "frequency", 5)
data.matrix$first_date<- as.factor(data.matrix$first_date)
```


#####Select only non-employee users
```{r}
data.matrix<-subset(data.matrix, data.matrix$emp_index=="N")
```


#####Removed add_type, cust_active, emp_index, date
```{r}
data.matrix <- cbind(data.matrix[,2],
                     data.matrix[,4:18],
                     data.matrix[,20:21],
                     data.matrix[,23:29]) 

dim(data.matrix)
names(data.matrix)
```


###Item-based Collaborative Filtering
#####Product Information
```{r}
data.recomendation <- data.matrix[,21:25]
```


#####Create item to item dataframe
```{r}
df.matrix <- matrix(NA, nrow=ncol(data.recomendation),
                    ncol=ncol(data.recomendation),
                    dimnames=list(names(data.recomendation),
                                  names(data.recomendation)))
```


#####matrix data frame
```{r}
recom.similarity <- as.data.frame(df.matrix)
```

#####cosine between two vectors function
```{r}
getCosine <- function(x,y) 
{
  this.cosine <- sum(x*y) / (sqrt(sum(x*x)) * sqrt(sum(y*y)))
  return(this.cosine)
}
```


#####cosine similarities on each product with each product
```{r}
for(i in 1:ncol(data.recomendation)) {
  for(j in 1:ncol(data.recomendation)) {
      recom.similarity[i,j]= getCosine(data.recomendation[i],data.recomendation[j])
  }
}

recom.similarity <- as.data.frame(recom.similarity)
recom.similarity
```


#####Get the top 3 neighbours for each product
#####Recommend only 2 product for each user for future
```{r}
product.neighbours <- matrix(NA, nrow=ncol(recom.similarity),
                          ncol=3,dimnames=list(colnames(recom.similarity)))
```


#####Creating a table of recomendation for each product
```{r}
for(k in 1:ncol(data.recomendation)) 
{
  product.neighbours[k,] <- (t(head(n=3,rownames(recom.similarity[order(recom.similarity[,k],decreasing=TRUE),][k]))))
}
product.neighbours
```


#####Adding recomended product to the particular user
```{r}
data.final<-cbind(data.matrix[,1],product.neighbours[,2],product.neighbours[,3])
colnames(data.final)<-c("CustomerID", "First_added_products", "Second_added_products")
```

#####First 20 Users with most likely buying products
```{r}
head(data.final, n=20)
```
